Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-30 intro.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The {\sl Napster} file-sharing was widely acknowledged as the 
``fastest growing Internet application ever'' in $2001$ when
$26$ million users shared $80$ million songs.
% File-sharing, originally introduced by
% {\sl Napster} (Section~\ref{section:background}), was widely
% acknowledged by $2001$ as the ``fastest growing Internet application ever" 
% with $26$ million users sharing $80$ million songs.
% % at the height of its popularity. 
Other file-sharing applications followed suit, 
including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
million concurrently-connected peers as well as 
{\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
\p\ telephony saw explosive growth with the advent of {\sl Skype}.
Since its introduction in $2003$,
{\sl Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers}; on March $5$th, $2012$,  
the astonishing figure of $35$ million concurrently-online users  
was reported~\cite{skypesymusers}.
%%
In addition, a startling number of diverse and successful
applications have been built based on a \p\ architecture. 
Some of them include:
distributed search engines~\cite{yaci}, 
distributed data-storage systems~\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging and dissemination applications~\cite{threedegrees,icpp08-pd}, 
event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 
naming services~\cite{cmm_chorddns_2002}, 
censor-resistant stores~\cite{cswh_freenet_2001} and
lately, even cloud-based platforms~\cite{mgpj_cloudsnap_2011}.

\p\ systems are implemented using \emph{overlay networks}. 
An overlay network, is a virtual system of nodes and logical links
created above an existing network; overlays provide
added value to the underlying layer by, for instance, offering QoS
content-distribution over the best effort Internet. 
Overlay nodes communicate through \emph{virtual connections} each 
of which may correspond to a path of possibly many physical links 
in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The single key issue that determines the efficiency of an overlay network,
is how well the overlay maps to the underlying network topology on which it
``rests''. 
Consider two nodes\footnote{In \p\ networks,
the participating nodes are typically user-PCs operating at the edge of the
Internet.} that are connected with each other via a path of overlay links.
If the application running on the nodes, generates heavy traffic along
the overlay path, it would be beneficial to
construct the overlay topology in a way that the number of 
underlying IP links between these two nodes is minimized.
% If the traffic patterns of the application
% are such that traffic flows heavily over the overlay link 
% between the two nodes,
% then it is beneficial to construct the overlay 
% topology such that the number of
% underlying IP links between these two nodes is minimized.
%%
Should the overlay network be constructed so that
it does not match the underlying topology well, 
the inherent \emph{topology mismatch} gives two
significant problems: first, the performance of the application per se, can be
adversely affected since traffic must flow over a larger,
redundant, number of physical hops resulting in poor user experience 
entailing noticeable latencies or jittering. 
%%
Second, other applications running
on the underlying network infrastructure are also adversely affected.
Studies have shown that highly popular \p\ applications contribute 
the largest portion of the overall 
Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
it was also projected that this traffic would reach
$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
Evidently, this constitutes a major burden for 
Internet Service Providers (ISPs) who must route 
all of this traffic to destinations at the edge of the Internet. 
If the \p\ overlay topology is poorly designed, 
the demand on the Internet's backbone infrastructure may 
substantially increase as traffic might have to flow
``back and forth'' several times between two neighboring ISPs
while trying to travel from the source to its destination node
in the overlay.
Hence, it is very critical that \p\ networks be laid out 
in ways that their topology matches, as closely as possible, the 
underlying IP topology.

For over a decade, researchers have extensively investigated  
various aspects of the topology mismatch problem.
The main objective of this paper is to offer a comprehensive 
survey of the work done in this and provide a
taxonomy of the proposed solutions. We point out synergies, as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select, amongst the different approaches, the one that is most suitable for
their particular application.
%%
The rest of the paper is organized as follows: 
Section~\ref{section:background} provides background on
overlay architectures including centralized, decentralized
unstructured and decentralized-structured \p\ systems. 
We also formally define the problem of topology mismatch 
% between an overlay and its underlying
% physical network 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Sections~\ref{section:unstructured} and~\ref{section:structured}
respectively outline the surveyed research efforts 
for decentralized-unstructured and decentralized-structured \p\ systems.
Conclusions can be found in Section~\ref{section:conclusion}.

>>>>

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The {\sl Napster} file-sharing system was widely acknowledged as the 
``fastest growing Internet application ever'' in $2001$ when it topped
$26$ million users sharing over $80$ million songs.
Other file-sharing applications followed suit, 
including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
million concurrently-connected peers, as well as 
{\sl BitTorrent} connecting over $150$ million monthly users by January $2012$~\cite{bittorrentusers}.
\p\ telephony saw explosive growth with the advent of {\sl Skype}.
Since its introduction in $2003$,
{\sl Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers} while on January $21$st, $2013$,
the astonishing figure of $50$ million concurrently-online users
was reported~\cite{skypesymusers}.
%%
Apart from the above, though, a startling number of diverse and successful
applications have been built based on \p\ architectures. Some of them include:
distributed search engines~\cite{yaci}, 
distributed data-storage systems~\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging and dissemination applications~\cite{threedegrees,icpp08-pd}, 
event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 
naming services~\cite{cmm_chorddns_2002}, 
censor-resistant stores~\cite{cswh_freenet_2001} and
lately, even cloud-based platforms~\cite{mgpj_cloudsnap_2011}.

% TODO SOME EXTRA REFERENCES IF NEEDED FOR THE USE OF OVERLAYS IN GENERAL
%
% IF ULTIMATELY DELETED CHECK TO REMOVE THE REFERENCES INCLUDED SINCE MOST
% PROBABLY ARE NOT USED ELSEWHERE.
%
%find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
%multicast}, as well as other proposals such as \emph{IntServ} and
%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
%largely due to the fact they require support from the IP
%infrastructure. On the other hand, an overlay network can be deployed on
%end-systems running the overlay protocol software, without the cooperation of ISPs.
%Academic research includes
%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

\p\ systems are implemented using \emph{overlay networks}. 
An overlay network, is a virtual system of nodes and logical links
created above an existing network; overlays provide an abstraction that
enables the implementaion of efficient, fully distributed, application-layer
services such as routing messages to destinations that are not known in advance
or offering QoS guarantees (i.e., in content-distribution) over best effort
infrastructures. Overlay nodes communicate through \emph{virtual connections}
each of which may correspond to a path of possibly many physical links 
in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The single key issue that determines the efficiency of an overlay network,
is how well the overlay maps to the underlying network topology on which it
``rests''. 
Consider two nodes\footnote{In \p\ networks,
the participating nodes are typically user-PCs operating at the edge of the
Internet.} that are connected with each other via a path of overlay links.
If the application running on the nodes, generates heavy traffic along
the overlay path, it would be beneficial to
construct the overlay topology in a way that the number of 
underlying IP links between these two nodes is minimized.
Should the overlay network be constructed so that
it does not match the underlying topology well, 
the inherent \emph{topology mismatch} gives two
significant problems. First, the performance of the application per se, can be
adversely affected since traffic must flow over a larger,
redundant, number of physical hops resulting in poor user experience 
entailing noticeable latencies or jittering. 
%%
Second, other applications running
on the underlying network infrastructure are also adversely affected.
Studies have shown that highly popular \p\ applications contribute 
the largest portion of the overall 
Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
it was also projected that this traffic would reach
$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
Evidently, this constitutes a major burden for 
Internet Service Providers (ISPs) who must route 
all of this traffic to destinations at the edge of the Internet. 
If the \p\ overlay topology is poorly designed, 
the demand on the Internet's backbone infrastructure may 
substantially increase as traffic might have to flow
``back and forth'' several times between two neighboring ISPs
while trying to travel from the source to its destination node
in the overlay.
Hence, it is very critical that \p\ networks be laid out 
in ways that their topology matches, as closely as possible, the 
underlying IP topology.

For over a decade, researchers have extensively investigated  
various aspects of the topology mismatch problem.
The main objective of this paper is to offer a comprehensive 
survey of the work done in this and provide a
taxonomy of the proposed solutions. We point out synergies, as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select, amongst the different approaches, the one that is most suitable for
their particular application.
%%
The rest of the paper is organized as follows: 
Section~\ref{section:background} provides background on
overlay architectures including centralized, decentralized-unstructured and
decentralized-structured \p\ systems.
We also formally define the problem of topology mismatch 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Sections~\ref{section:unstructured} and~\ref{section:structured}
respectively outline the surveyed research efforts 
for decentralized-unstructured and decentralized-structured \p\ systems.
Conclusions can be found in Section~\ref{section:conclusion}.

### 2014-01-30 mard-survey.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

\begin{abstract}
\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.  %% no commas here
They are implemented via an overlay network abstraction, on top 
of the best-effort IP infrastructure of the Internet. 
\p\ systems offer a plethora of desirable features to 
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, to name just a few.
Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
applications from leveraging these features.
One major weakness, identified early on, is the \emph{topology mismatch problem}
between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability and efficiency of
applications built atop.
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the suggested solutions
by discussing their novelty, advantages and weaknesses.
Finally, we offer an analysis as well as pictorial comparisons 
of the reviewed approaches, since we aim to offer a comprehensive
reference for developers, system architects and researchers in the field.

@@

\keywords{Distributed systems, peer-to-peer architectures, overlay network,
topology awareness}

@@

% \item Proximity Neighbor Selection\\
%Finally, the third approach, constructs the routing tables using proximity
%knowledge. Tapestry and Pastry's mechanisms of routing table
%maintenance try to minimize the distance to nodes appearing in a peer's routing
%table. Since routing is based on longest node ID prefix match, messages 
%are gradually forwarded to nearby nodes at each routing step.
%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
%neighbor selection in order to create a scheme that is (more) location-aware
%compared to the other well-known DHTs (CAN, Chord).
%
%\end{itemize}
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
%selection (topology selection, cluster, cache etc), supports dynamic update,
%runtimes}}
%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
%
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
%
%{\sethlcolor{yellow}\hl{
%HA: maybe add also the year of publication, and see if there is a pattern in
%terms of the method and the year??}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
%       ATTENTION: NEEDS REWORDING
%
% P4P [16],instead, aimed to provide a general framework for P2P
%applications by developing ISP-application interfaces so that
%application software can use the network layer information,
%such as congestion status and network topology, for better
%performance and resource utilization.

>>>>

\begin{abstract}
\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.
They are implemented via an overlay network abstraction, on top 
of the Internet's best-effort IP infrastructure. 
\p\ systems support a plethora of desirable features to
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, to name just a few.
Unfortunately, inherent weaknesses of early deployments of \p\ systems, prevented 
applications from leveraging the full potential of the paradigm.
One major weakness, identified early on, is the \emph{topology mismatch problem}
between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability and efficiency of
applications built atop.
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the suggested solutions
by discussing their novelty, advantages and weaknesses.
Finally, we offer an analysis as well as pictorial comparisons 
of the reviewed approaches, since we aim to offer a comprehensive
reference for developers, system architects and researchers in the field.

@@

\keywords{Distributed systems, peer-to-peer architectures, p2p, overlay network,
topology awareness}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: SOME INFORMATION ON ISP-ASSISTED TOPOLOGY AWARENESS
%
% IN CASE REVIEWERS POINT OUT THAT WE SHOULD ADD THOSE AS WELL.
%
%       ATTENTION: NEEDS REWORDING
%
% P4P [16],instead, aimed to provide a general framework for P2P
%applications by developing ISP-application interfaces so that
%application software can use the network layer information,
%such as congestion status and network topology, for better
%performance and resource utilization.

### 2014-01-30 structured.tex
2013.05.20 (Vassilis) Minor changes to abstract, introduction, background and unstructured (tex, plot and tables) sections. Useless comment clean-up. Various rephrasing on some reviewed content. Added references to the tables. Some image revisited. Discussion text added explaining the fact that many of the papers combine multiple methodologies.

Maintaining several host states at different layers, makes any content migration
costly. Additionally, the method does not make any continuing effort to remap
the overlay structure after a node successfully joins in order to adapt its
state to any condition change. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
traditional DHTs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Mithos}
\emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
incremental probing to find near optimal node placement, and is classified by
its authors as an integration of geographic layout and proximity routing overlay
optimization methods.

The bootstrap phase of Mithos starts with a subset of existing members as the
first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. To
avoid a local minima, Mithos probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

@@

>>>>

Maintainece of several host states at different layers, renders any content
migration a costly process. Additionally, the method does not make any
continuing effort to remap
the overlay structure after nodes successfully join resulting into poor
adaptation to changing conditions. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
traditional DHTs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Mithos}
\emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
incremental probing to find near optimal node placement, and is an integration
of geographic layout and proximity routing overlay optimization methods.

The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
the first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. To
avoid a local minima, Mithos probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

@@

% \item Proximity Neighbor Selection\\
%Finally, the third approach, constructs the routing tables using proximity
%knowledge. Tapestry and Pastry's mechanisms of routing table
%maintenance try to minimize the distance to nodes appearing in a peer's routing
%table. Since routing is based on longest node ID prefix match, messages
%are gradually forwarded to nearby nodes at each routing step.
%\cite{castro_proximityp2p_2002} argues on how Pastry exploits proximity
%neighbor selection in order to create a scheme that is (more) location-aware
%compared to the other well-known DHTs (CAN, Chord).
%
%\end{itemize}
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria: based on which protocol (eg Gnutella, CAN etc), peer
%selection (topology selection, cluster, cache etc), supports dynamic update,
%runtimes}}
%%\textbf{Algorithm} & \textbf{Overlay structure} & \textbf{Forwarding} &
%\textbf{Cache} & \textbf{Overlay optimization} & \textbf{Proximity information}
%& \textbf{Base protocol} & \textbf{Dynamic update} & \textbf{Runtime} \\
%
%
%{\sethlcolor{yellow}\hl{
%HA: Possible criteria2: Overlay optimization structure, base protocol (eg
%Gnutella, CAN etc), dynamic update, runtimes, scalability}}
%
%{\sethlcolor{yellow}\hl{
%HA: maybe add also the year of publication, and see if there is a pattern in
%terms of the method and the year??}}
